Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2025-02-17 ms.qmd
Fixing up requested changes

Contrary to acclimation capacity variance in physiological rates increased across habitats with effect sizes being 3-5 times larger than those observed for acclimation of mean trait values. Mechanistically, it is unclear what exactly is contributing to the increased variation in physiological rates at higher temperatures, but it is likely the result of increased among-individual variability in how biochemical, cellular and physiological processes function at higher temperatures to maintain homeostasis [@tattersall2012coping; @angilletta2009thermal; @somero1995proteins; @fields2001protein; @schulte2011thermal]. Higher temperatures increase membrane fluidity affecting electrochemical gradients and impacting protein structure and function [@somero1995proteins; @fields2001protein; @tattersall2012coping]. Such challenges (among others) may expose among-individual variation within a population. Indeed, there is considerable variation in acclimation capacity among individuals which would increase variance in thermal performance curves within populations [@loughland2020differences; @schulte2011thermal].

@@

Using $\Delta lnRR_{Q_{10}}$ for each trait within a study we fit a model that estimated the meta-anlaytic mean effect size for each habitat (freshwater, marine and terrestrial) accounting for acclimation time and maximum temperature as in our main analysis, along with random effects of study, species, phylogeny, and trait. Our analyses given quantitatively and qualitatively similar resuls to our main analysis @fig-delta. Acclimation time did not explain variation in $\Delta lnRR_{Q_{10}}$ (slope = `r mean(b_c_accl_time_slope)`, 95\% CI: `r quantile(b_c_accl_time_slope, c(0.025, 0.975))`, p = `r pmcmc(b_c_accl_time_slope)`) and there was weak evidence that maximum temperature explained variation in $\Delta lnRR_{Q_{10}}$ (slope = `r mean(b_max_temp_c)`, 95\% CI: `r quantile(b_max_temp_c, c(0.025, 0.975))`, p = `r pmcmc(b_max_temp_c)`).

>>>>

Contrary to acclimation capacity, variance in physiological rates increased across habitats with effect sizes being 3-5 times larger than those observed for acclimation of mean trait values. Mechanistically, it is unclear what exactly is contributing to the increased variation in physiological rates at higher temperatures, but it is likely the result of increased among-individual variability in how biochemical, cellular and physiological processes function at higher temperatures to maintain homeostasis [@tattersall2012coping; @angilletta2009thermal; @somero1995proteins; @fields2001protein; @schulte2011thermal]. Higher temperatures increase membrane fluidity affecting electrochemical gradients and impacting protein structure and function [@somero1995proteins; @fields2001protein; @tattersall2012coping]. Such challenges (among others) may expose among-individual variation within a population. Indeed, there is considerable variation in acclimation capacity among individuals which would increase variance in thermal performance curves within populations [@loughland2020differences; @schulte2011thermal].

@@

Using $\Delta lnRR_{Q_{10}}$ for each trait within a study we fit a model that estimated the meta-anlaytic mean effect size for each habitat (freshwater, marine and terrestrial) accounting for acclimation time and maximum temperature as in our main analysis, along with random effects of study, species, phylogeny, and trait. Our analyses give quantitatively and qualitatively similar results to our main analysis (@fig-delta). Acclimation time did not explain variation in $\Delta lnRR_{Q_{10}}$ (slope = `r mean(b_c_accl_time_slope)`, 95\% CI: `r quantile(b_c_accl_time_slope, c(0.025, 0.975))`, p = `r pmcmc(b_c_accl_time_slope)`) and there was weak evidence that maximum temperature explained variation in $\Delta lnRR_{Q_{10}}$ (slope = `r mean(b_max_temp_c)`, 95\% CI: `r quantile(b_max_temp_c, c(0.025, 0.975))`, p = `r pmcmc(b_max_temp_c)`).
